A free/open-source Kazakh-Tatar machine translation system

### 2013-04-22 paper/paper.tex
add › to hard/soft sign

Within the Apertium project, work on MT systems between Turkic languages has been started (Turkish--Kyrgyz, Azeri--Turkish, Turkish-Bashkir), but the Kazakh--Tatar system described by the present study is the closest to production-ready of them.  Among these systems is a prototype Tatar-Bashkir machine translation system which was built by the authors of this paper \citep{tyerswashingtonsalimzyanbattalov12}; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage.

@@

While Tatar and Kazakh both have a velar/uvular obstruent distinction (e.g., /k/ vs.\ /q/) that interacts with 
adjacent vowels, the Tatar orthography only has one series of letters (e.g., ‹к›), relying on adjacent 
vowels (and employing ъ `hard sign' and ь `soft sign' when these fail) to differentiate the two, and Kazakh 
has two series of obstruents (e.g., ‹к› and ‹қ›).

>>>>

Within the Apertium project, work on MT systems between Turkic languages has been started (Turkish--Kyrgyz, Azeri--Turkish, Tatar--Bashkir), but the Kazakh--Tatar system described by the present study is the closest to production-ready of them.  Among these systems is a prototype Tatar--Bashkir machine translation system which was built by the authors of this paper \citep{tyerswashingtonsalimzyanbattalov12}; due to the closeness of these languages, it proved to provide high accuracy in its translations, but being a prototype system by design, had relatively low coverage.

@@

While Tatar and Kazakh both have a velar/uvular obstruent distinction (e.g., /k/ vs.\ /q/) that interacts with 
adjacent vowels, the Tatar orthography only has one series of letters (e.g., ‹к›), relying on adjacent 
vowels (and employing ‹ъ› `hard sign' and ‹ь› `soft sign' when these fail) to differentiate the two, and Kazakh 
has two series of obstruents (e.g., ‹к› and ‹қ›).
